Empirical Study of Scientific Computer Software: Objectives and Research Questions

### 2020-04-28 OverallResearchProposal/ObjectivesAndResearchQuestions.tex
Edits to address comments by Ao

In the above definition, software is considered to be an application (app)
and/or a library.  The outcome of the development process is not just software.
It is \{software + artifacts\}.  \wss{softwartifacts?} \ad{try softifacts?}The artifacts consist of
such work products as documentation, build scripts, code, test cases,
contributor's guides, etc.  The quality of sustainability cannot be assessed by
looking at the software alone, or the code alone; it depends on everything,
including the processes and methodologies used \ad{during the development?}.  Further information on the
relevant definitions and context is given in the next section.

The notion of ``reasonable amount of time and energy'' is part of the objective,
but reasonable is not defined in the objective.  What is reasonable is going to
depend on the context.  Problems that are large, complex and/or critical are
going to have a greater ``budget'' of time and energy.  When comparing different
alternatives for processes and methodologies, preference would be given to the
alternative that costs less time and energy, assuming that this alternative
achieves the same \ad{or higher} level of sustainability as its competition.  The assumptions
listed in the next section outline the typical space of problems where it makes
sense to aim for sustainable \{software + artifacts\}.

@@

The objective is not phrased to target a specific software domain, but one of
the research assumptions is that we will focus on Scientific Computing Software
(SCS) examples and applications.  This focus is because creating sustainable
software+artifacts is more feasible when the software domain is well understood,
which is the case for SCS.  Moreover, SCS is an important, often neglected,
domain of software.

@@

The overall objective is too ambiguous as given.  It requires additional
information/interpretation.  Therefore this section provides definitions of the
relevant terms and a list of the assumptions that define the potential contexts
where sustainable software+artifact development is reasonable.

@@

  This definition is general; it is not specific to software and it is silent on
  the specific needs of the present and the future.  The definition can be made
  less abstract by introducing the context of software+artifacts and thinking
  separately about the needs of the present and the future.  For the needs of
  the present, the software should be correct, reliable and usable. For the
  needs of the future, given uncertainty, this means supporting change from
  relevant concerns (political, economic, social, technical, legal and
  environmental) should be possible with a reasonable level of resources.
  Supporting the future then means producing software+artifacts that are
  maintainable, portable, reusable and reproducible.

@@

  \emph{Sustainable software means software+artifacts that are correct,
    reliable, and usable (for the present) and maintainable, portable, reusable
    and reproducible (for the future).}

@@

\item[Software] Application and/or library.
\item[Application (App)] An executable program. \wss{needs a
    proper definition, with citation(s)}

``Application software is a program or group of programs designed for end users.
These programs are divided into two classes: system software and application
software. While system software consists of low-level programs that interact
with computers at a basic level, application software resides above system
software and includes applications such as database programs, word processors
and spreadsheets. Application software may be bundled with system software or
published alone.''\citep{TechopediaApplication2018}\ad{Definition from
Techopedia.com}

@@

Producing sustainable software+artifacts is not a trivial undertaking.  Not
every project needs to aim for sustainability.  Our position is that the time
and energy is justified when one or more of the following assumptions applies.

\begin{itemize}
\item The software project is going to be long lived, where long life means at
  least 10 years.
\item The software+artifacts should interest multiple stakeholders (not just the
  original developer), with different interests and backgrounds.
\item The software is safety relevant, such as software for nuclear safety,
  medical imaging or computational medicine.
\item The software explicitly, or implicitly, is part of a program family of
  related software.  When the program family assumption means that there are
  related programs that are less effort to design and build together than to
  design and build as separate projects.  Program family development depends on
  satisfying three hypotheses \citep{Weiss1997}:
\begin{itemize}
\item Redevelopment hypothesis – most software development involved in producing
  an individual family members should be redevelopment because the family
  members have so much in common.
\item Oracle Hypothesis – the changes that are likely to occur during the soft-
  ware’s \ad{software's} lifetime are predictable.
\item Organizational Hypothesis – designers can organize the software and the
  development effort so that predicted changes can be made independently.
\end{itemize}
\item The project recognizes that software artifacts other than the code are
  relevant.
\item The domain of the software project is well understood.  This assumption is
  related to the oracle hypothesis.
\end{itemize}

@@

A good starting point for understanding how to develop sustainable SCS software+artifacts is
to look at how SCS is currently developed.  By measuring relevant qualities and
relating them to development practices, we can form an idea of what is currently
working well.

@@

\begin{enumerate}
\item What knowledge is currently captured in the artifacts generated by
  existing open source SCS projects?
\item What principles, processes, methodologies and tools are used by existing
  open source SCS projects?
\item What are the ``pain points'' for developers working on SCS projects?  What
  aspects of the existing processes, methodologies and tools do they consider
  could potentially be improved?
\item For a given software product (software+artifacts), how can we produce a
  reasonable measure of the software qualities (correctness, reliability,
  usability etc.) with a few hours of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of usability with a few days worth of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of maintainability (with respect to likely changes) with a
  few days worth of effort?
\item For a given software product (software+artifacts), how can we produce a
  meaningful measure of reproducibility/replicability with a few days worth of
  effort?
\item With a reasonable effort and time, how can we compare the productivity
  between different processes, methodologies and tools?
\item Can a correlation be identified between software projects that use best \ad{better}
  practices for the development process and methodologies and improvement in the
  qualities of usability, performance and reproducibility/replicability?
\ad{Change "best" to "better"? The best one isn't likely to be found, and better
practices may also result in improvements.}
\end{enumerate}

@@

>>>>

In the above definition, software is considered to be an application (app)
and/or a library.  The outcome of the development process is not just software.
It is \{software + artifacts\}.  From this point onward, we will use the term
softifacts when we are referencing the combination of software and the
associated artifacts.  The artifacts consist of such work products as
documentation, build scripts, code, test cases, contributor's guides, etc.  The
quality of sustainability cannot be assessed by looking at the software alone,
or the code alone; it depends on everything, including the processes and
methodologies used during the development.  Further information on the relevant
definitions and context is given in the next section.

The notion of ``reasonable amount of time and energy'' is part of the objective,
but reasonable is not defined in the objective.  What is reasonable is going to
depend on the context.  Problems that are large, complex and/or critical are
going to have a greater ``budget'' of time and energy.  When comparing different
alternatives for processes and methodologies, preference would be given to the
alternative that costs less time and energy, assuming that this alternative
achieves the same (or a higher) level of sustainability as its competition.  The
assumptions listed in the next section outline the typical space of problems
where it makes sense to aim for sustainable softifacts.

@@

The objective is not phrased to target a specific software domain, but one of
the research assumptions is that we will focus on Scientific Computing Software
(SCS) examples and applications.  This focus is because creating sustainable
softifacts is more feasible when the software domain is well understood,
which is the case for SCS.  Moreover, SCS is an important, often neglected,
domain of software.

@@

The overall objective is too ambiguous as given.  It requires additional
information/interpretation.  Therefore this section provides definitions of the
relevant terms and a list of the assumptions that define the potential contexts
where sustainable softifacts development is reasonable.

@@

  This definition is general; it is not specific to software and it is silent on
  the specific needs of the present and the future.  The definition can be made
  less abstract by introducing the context of softifacts and thinking
  separately about the needs of the present and the future.  For the needs of
  the present, the software should be correct, reliable and usable. For the
  needs of the future, given uncertainty, this means supporting change from
  relevant concerns (political, economic, social, technical, legal and
  environmental) should be possible with a reasonable level of resources.
  Supporting the future then means producing softifacts that are
  maintainable, portable, reusable and reproducible.

@@

  \emph{Sustainable software means softifacts that are correct,
    reliable, and usable (for the present) and maintainable, portable, reusable
    and reproducible (for the future).}

@@

\item[Software] Application and/or library.
\item[Application (App)] ``Software designed to fulfill specific needs of a
  user.'' \citep{IEEEStdGlossarySET1990}

``Application software is a program or group of programs designed for end users.
These programs are divided into two classes: system software and application
software. While system software consists of low-level programs that interact
with computers at a basic level, application software resides above system
software and includes applications such as database programs, word processors
and spreadsheets. Application software may be bundled with system software or
published alone.''\citep{TechopediaApplication2018} \ad{Definition from
Techopedia.com}

@@

Producing sustainable softifacts is not a trivial undertaking.  Not
every project needs to aim for sustainability.  Our position is that the time
and energy is justified when one or more of the following assumptions applies.

\begin{itemize}
\item The software project is going to be long lived, where long life means at
  least 10 years.
\item The softifacts should interest multiple stakeholders (not just the
  original developer), with different interests and backgrounds.
\item The software is safety relevant, such as software for nuclear safety,
  medical imaging or computational medicine.
\item The software explicitly, or implicitly, is part of a program family of
  related software.  When the program family assumption means that there are
  related programs that are less effort to design and build together than to
  design and build as separate projects.  Program family development depends on
  satisfying three hypotheses \citep{Weiss1997}:
\begin{itemize}
\item Redevelopment hypothesis – most software development involved in producing
  an individual family members should be redevelopment because the family
  members have so much in common.
\item Oracle Hypothesis – the changes that are likely to occur during the soft-
  ware’s \ad{software's} lifetime are predictable.
\item Organizational Hypothesis – designers can organize the software and the
  development effort so that predicted changes can be made independently.
\end{itemize}
\item The project recognizes that software artifacts other than the code are
  relevant.
\item The domain of the software project is well understood.  This assumption is
  related to the oracle hypothesis.
\end{itemize}

@@

A good starting point for understanding how to develop sustainable SCS softifacts is
to look at how SCS is currently developed.  By measuring relevant qualities and
relating them to development practices, we can form an idea of what is currently
working well.

@@

Define what is meant by meaningful (what criteria need to be satisfied), since
this term is used several times in the research questions.

\begin{enumerate}
\item What knowledge is currently captured in the artifacts generated by
  existing open source SCS projects?
\item What principles, processes, methodologies and tools are used by existing
  open source SCS projects?
\item What are the ``pain points'' for developers working on SCS projects?  What
  aspects of the existing processes, methodologies and tools do they consider
  could potentially be improved?
\item For a given software product (softifacts), how can we produce a
  reasonable measure of the software qualities (correctness, reliability,
  usability etc.) with a few hours of effort?
\item For a given software product (softifacts), how can we produce a
  meaningful measure of usability with a few days worth of effort?
\item For a given software product (softifacts), how can we produce a
  meaningful measure of maintainability (with respect to likely changes) with a
  few days worth of effort?
\item For a given software product (softifacts), how can we produce a
  meaningful measure of reproducibility/replicability with a few days worth of
  effort?
\item With a reasonable effort and time, how can we compare the productivity
  between different processes, methodologies and tools?
\item Can a correlation be identified between software projects that use best 
  practices for the development process and methodologies and improvement in the
  qualities of usability, performance and reproducibility/replicability?
\end{enumerate}

@@

NOTES

- what is needed to produce softifacts sustainably?

- measuring process aimed at 'producing softifacts sustainably'. part of it could be measuring effort.
